---
title: "gusto_les_cleaning"
author: "Fran Querdasi"
date: "11/9/2020"
output: 
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
---

#Setup for knitting
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

## Load libraries

```{r}
library(tidyverse)
library(knitr)
library(readr)
library(psych)
library(lubridate)
library(readxl)
library(stringr)
```


## Import data
```{r}
les_file <- "../../data/stress/GUSTO_Y04_Life Experiences Survey_20190620.xlsx"

les_data <-
  read_excel(les_file)
```

## Compute total events endorsed (les_events_total)
```{r}
#get LES tally vars:
les_tally_vars <-
  les_data %>% 
  select(matches("^q([0-9]|[1-4][0-9]|50)$")) %>% 
  names()

#account for questions that have sub-questions
les_extra_tally_vars <-
  les_data %>% 
  select(matches("^q(5[a-g]|15[a-h])$")) %>% 
  names()

les_total_tally_vars = c(les_tally_vars, les_extra_tally_vars)

#Compute total number of events (regardless of time)

#relabel all of the 9997 and -9999 to NA (data dict states that 9997 - NA - not sure what -9999 is then?)
les_data <-
  les_data %>% 
  mutate_at(vars(les_total_tally_vars), list(~ ifelse( . == 9997, NA, .)))

les_data <-
  les_data %>% 
  mutate_at(vars(les_total_tally_vars), list(~ ifelse( . == -9999, NA, .)))

#relabel all of the values in the tally vars to be "1" if the event has been endorsed (2 means endorsed 7 mo to 1 yr ago, 1 means 0 to 6mo ago)
les_data <-
  les_data %>% 
  mutate_at(vars(les_total_tally_vars), list(~ ifelse( . == 2, 1, .)))

#account for places where a rating was given but event was not listed as endorsed:

#get column numbers for question variables
q_col_nums_1 <- grep("^q([0-9]|[1-4][0-9]|50)$", colnames(les_data))
q_col_nums_2 <- grep("^q(5[a-g]|15[a-h])$", colnames(les_data))
q_col_nums <- c(q_col_nums_1, q_col_nums_2)

#convert question variables to double for the for loop below to work
les_data <-
  les_data %>% 
  mutate_at(vars(les_total_tally_vars), .funs = list(~as.double(.)))

#If rating variable is not NA, assign corresponding question variable a value of 1
for (rownum in 1:nrow(les_data)) {
  for (idx in 1:length(q_col_nums)) {
  if (!is.na(les_data[rownum, q_col_nums[idx]+4])) { #corresponding rating variables are always 4 columns after q var
    les_data[rownum, q_col_nums[idx]] = 1
  }
  }
}

#select dataframe containing tally vars
les_test <-
  les_data %>% 
  select(all_of(les_total_tally_vars))

#now use the fact that all non-endorsed events have NA values and all endorsed values are scored 1 to sum all endorsed events
les_data$les_events_total <- rowSums(les_test, na.rm = TRUE)

```

## Compute separate total for events 0-6mo ago and 7-12mo ago (les_sixmnth_total & les_oneyr_total) 
```{r}
#0 to 6 months
les_sixmo_vars <-
  les_data %>% 
  select(contains("0to6mths")) %>% 
  names()

les_data <-
  les_data %>% 
  mutate_at(vars(les_sixmo_vars), as.character()) #coerce to character so that the ifelse below will work

les_data <-
  les_data %>% 
  mutate_at(vars(les_sixmo_vars), list(~ ifelse( . == "1_yes", 1, .)))

les_sixmo_data <-
  les_data %>% 
  select(all_of(les_sixmo_vars))

les_data$les_sixmnth_total <- rowSums(les_sixmo_data, na.rm = TRUE)

#7 months to 1 yr
les_oneyr_vars <-
  les_data %>% 
  select(contains("7mthto1yr")) %>% 
  names()

les_data <-
  les_data %>% 
  mutate_at(vars(les_oneyr_vars), as.character()) #coerce to character so that the ifelse below will work

les_data <-
  les_data %>% 
  mutate_at(vars(les_oneyr_vars), list(~ ifelse( . == "1_yes", 1, .)))

les_oneyr_data <-
  les_data %>% 
  select(all_of(les_oneyr_vars))

les_data$les_oneyr_total <- rowSums(les_oneyr_data, na.rm = TRUE)
```

## Compute total N events rated positively
```{r}
les_rating_vars <-
  les_data %>% 
  select(contains("rating")) %>% 
  names()

```

## Compute a sum score
```{r}
#this will be compared to the score calculated by GUSTO, lestotrating_y4
les_rating_data <-
  les_data %>% 
  select(all_of(les_rating_vars)) %>% 
  select(-total_rating) #this is GUSTO's score and if it is not eliminated from these vars, it will be included in row sum

les_data$les_score <- rowSums(les_rating_data, na.rm = TRUE)
```

## Compute a sum of all positive scores (les_pos_score)
```{r}
les_pos_data <-
  les_rating_data %>% 
  mutate_all(., list(~ ifelse( . <= 0, NA, .))) #create copy of these vars that only preserves positive ratings (all neg and neutral ratings are changed to NA)

les_data$les_pos_score <- rowSums(les_pos_data, na.rm = TRUE)
```

## Compute a sum of all negative scores (les_neg_score)
```{r}
les_neg_data <-
  les_rating_data %>% 
  mutate_all(., list(~ ifelse( . >= 0, NA, .))) #create copy of these vars that only preserves negative ratings (all pos and neutral ratings are changed to NA))

les_data$les_neg_score <- rowSums(les_neg_data, na.rm = TRUE)
```

## Compute a sum of all 'child' events (les_child_events_total)
```{r}
les_child_vars <-
  les_data %>% 
  select('q2', 'q7', 'q15g', 'q19', 'q20', 'q26', 'q35', 'q37') %>% 
  names()

les_child_data <-
  les_data %>% 
  select(all_of(les_child_vars))

les_data$les_child_events_total <- rowSums(les_child_data, na.rm = TRUE)

```

## Compute total score on 'child' events
```{r}
les_child_score_data <-
  les_data %>% 
  select(all_of(les_rating_vars)) %>% 
  select(matches("^q(2|7|15g|19|20|26|35|37)_rating$"))

les_data$les_child_score <- rowSums(les_child_score_data, na.rm = TRUE)
les_child_score_data$les_child_score <- rowSums(les_child_score_data, na.rm=TRUE)

les_check <-
  les_data %>% 
  select(
    PSCID,
    les_child_score
  )

les_child_score_data <-
  les_child_score_data %>% 
  left_join(
    les_check, by = "les_child_score"
  )
```

## Compute tally of positive "child" events
```{r}
#select from les_pos_data only those that are child events
les_child_pos_data <-
  les_pos_data %>% 
  select(matches("^q(2|7|15g|19|20|26|35|37)_rating$")) %>% 
  mutate_all(., list(~ ifelse( . > 0, 1, .)))

les_data$les_child_pos_events_total <- rowSums(les_child_pos_data > 0, na.rm = TRUE)
```

## Compute tally of negative "child" events
```{r}
#select from les_pos_data only those that are child events
les_child_neg_data <-
  les_neg_data %>% 
  select(matches("^q(2|7|15g|19|20|26|35|37)_rating$")) %>% 
  mutate_all(., list(~ ifelse( . < 0, 1, .)))

les_data$les_child_neg_events_total <- rowSums(les_child_neg_data, na.rm = TRUE)
```

## Quick descriptive stats
```{r}
les_data %>% 
  ggplot(aes(les_events_total)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LES_events_hist.png",
  width = 10,
  height = 10
  )

les_data %>% 
  ggplot(aes(les_score)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LES_score_hist.png",
  width = 10,
  height = 10
  )

#remove obvious outliers
les_data %>% 
  filter(les_score < 25 & les_score > -50) %>% 
  ggplot(aes(les_score)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LES_score_rmoutliers_hist.png",
  width = 10,
  height = 10
  )

les_data %>% 
  ggplot(aes(les_pos_score)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LES_pos_score_hist.png",
  width = 10,
  height = 10
  )

les_data %>% 
  ggplot(aes(les_neg_score)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LES_neg_score_hist.png",
  width = 10,
  height = 10
  )
```

## Create cleaned dataset
```{r}
les_data_clean <-
  les_data %>% 
  select(
    subjid = PSCID,
    Date_taken,
    person_filling,
    les_events_total,
    les_sixmnth_total,
    les_oneyr_total,
    les_score,
    les_pos_score,
    les_neg_score,
    les_child_events_total,
    les_child_score,
    les_child_pos_events_total,
    les_child_neg_events_total
  )

#determine which year the data is from, extract that part of the file name
year <- substr(substring(les_file, regexpr("(_[M|Y][0-9]?[1-9]_)", les_file)), 1, 4)

#change column names
colnames(les_data_clean) <- paste0(colnames(les_data_clean), year)
#change first column name back (because ID is constant across age and needs to be the same so we can merge files)
colnames(les_data_clean)[1] <- str_replace(colnames(les_data_clean)[1], year, "")
```

## Write output file
```{r}
#create file name which contains today's date and child age data were collected in years
out_file_name <- str_c("../../data/stress/GUSTO", year, "_LES_clean_", str_replace_all(Sys.Date(), "-", ""), ".csv")

#write output file
write_csv(les_data_clean, out_file_name)
```

