---
title: "gusto_leq_cleaning"
author: "Fran Querdasi"
date: "11/22/2020"
output: 
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
---

#Setup for knitting
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

## Load libraries

```{r}
library(tidyverse)
library(knitr)
library(readr)
library(psych)
library(lubridate)
library(readxl)
```

## Year 4 LEQ

### Import data
```{r}
leq_file <- "../../data/stress/GUSTO_Y04_Life Events Questionnaire_20190620_updatedwage.xlsx"

leq_data <-
  read_excel(leq_file)

```

### Find total number of life events endorsed by each participant - leq_events_total
```{r}
#Find total number of life events endorsed for each participant

#1. Select all variable names that have: [1-10]_y4 within the  name
leq_tally_vars <-
  leq_data %>% 
  select(matches("([1-9]|10)_y4")) %>% 
  names()

#2. Update tally variables to be 0 or 1 corresponding to "0_no" or "1_yes" on leq_tally_vars
leq_data_cleaning <-
  leq_data %>% 
  mutate_at(vars(leq_tally_vars), list(~ ifelse( . == '0_no', 0, 1)))

#3. Create variable that is sum across all tally vars
leq_events_total <-
  leq_data_cleaning %>% 
  select(all_of(leq_tally_vars)) %>% 
  rowSums(na.rm = TRUE)

leq_data_cleaning <- cbind(leq_data_cleaning, leq_events_total)
```

###Find total number of events rated as "negative" - leq_neg_events_total
```{r}
#1. Select all variables that contain pos/neg rating
leq_rating_vars <-
  leq_data %>% 
  select(contains("exp")) %>% 
  names()


#2. Get a sum of events classified as "negative" for each row
leq_test <-
  leq_data_cleaning %>% 
  select(all_of(leq_rating_vars))

leq_data_cleaning$leq_neg_events_total <- rowSums(leq_test == "negative", na.rm = TRUE)

```

###Find total number of events rated as "positive" - leq_pos_events_total
```{r}

leq_data_cleaning$leq_pos_events_total <- rowSums(leq_test == "positive", na.rm = TRUE)
```

### Find total number of events not classified - leq_notrated_total **not working**
```{r}
#1. change ambiguous notation - change all -9999, which means no rating was given, to "notrated"
leq_data_cleaning <-
  leq_data_cleaning %>% 
  mutate_at(vars(leq_rating_vars), list(~ ifelse( . == '-9999', 'notrated', .)))

leq_test <-
  leq_data_cleaning %>% 
  select(leq_rating_vars) 

leq_test <-
  leq_test %>% 
  mutate_all(as.character()) 

#leq_data_cleaning$leq_notrated_total <- rowSums(leq_test == "notrated", na.rm = T) #not sure why this is not working

leq_data_cleaning <-
  leq_data_cleaning %>% 
  mutate(
    leq_notrated_total = leq_events_total - (leq_pos_events_total + leq_neg_events_total)
  )

```

### Find approximate child age at each endorsed event **make age bin vars?**
```{r}
#get list of variables
leq_age_vars <-
  leq_data %>% 
  select(contains("age")) %>% 
  names()

#change -999 values to NA
leq_data_cleaning <-
  leq_data_cleaning %>% 
  mutate_at(vars(leq_age_vars), list(~ ifelse( . == '-9999', NA, .)))

#convert age in days to age in months
days_to_months <- function(x) {
  x = x / 30.4167
  }

leq_data_cleaning <-
  leq_data_cleaning %>% 
  mutate_at(vars(leq_age_vars), list(~days_to_months(.)))

#make age bin variables? 
```
### Quick descriptive stats

```{r}
#summarise variables
leq_data_cleaning %>% 
 summarise_at(
    vars(
      leq_events_total,
      leq_neg_events_total,
      leq_pos_events_total
      ),
    funs(mean, sd, min, max), na.rm = TRUE
  )

#avg event is 1.5, max is 6 (scale max is 15), avg more pos_events than stressful events

#histograms of variables
leq_data_cleaning %>% 
  ggplot(aes(leq_events_total)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LEQ_events_hist.png",
  width = 10,
  height = 10
  )

leq_data_cleaning %>% 
  ggplot(aes(leq_neg_events_total)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LEQ_neg_eventss_hist.png",
  width = 10,
  height = 10
  )

leq_data_cleaning %>% 
  ggplot(aes(leq_pos_events_total)) +
  geom_histogram(binwidth = 1)

ggsave(
  "../../figures/descriptives/GUSTO_Y04_LEQ_pos_events_hist.png",
  width = 10,
  height = 10
  )

#all variables positively skewed 
```


### Select variables of interest

```{r}
leq_data_clean <-
  leq_data_cleaning %>%
  select(
    subjid,
    leq_age_vars,
    leq_events_total,
    leq_pos_events_total,
    leq_neg_events_total
  )

#determine which year the data is from, extract that part of the file name
year <- substr(substring(les_file, regexpr("(_[M|Y][0-9]?[1-9]_)", leq_file)), 1, 4)

#columns already have year in them so just change the last ones
ind_l <- length(colnames(leq_data_clean))-2
ind_h <- length(colnames(leq_data_clean))
colnames(leq_data_clean)[ind_l:ind_h] <- paste0(colnames(leq_data_clean)[ind_l:ind_h], year)
```
### Write output file
```{r}
#create file name which contains today's date and child age data were collected in years
out_file_name <- str_c("../../data/stress/GUSTO", year, "_LEQ_clean_", str_replace_all(Sys.Date(), "-", ""), ".csv")

write_csv(leq_data_clean, out_file_name)
```

